RProtoBuf: Efficient Cross-Language Data Serialization in R

### 2014-01-09 papers/rjournal/eddelbuettel-stokely.Rnw
Fix some typos / add misisng articles in Jereoen's new OpenCPU section. Move this example before the MapReduce one.
It uses the serialize_pb method but we haven't described that here.  A
short introduction of that concept is needed for the application.

I think some of the first paragraph here should be moved to the
introduction so this section can focus only on an application.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@726 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Application: Sending/receiving Interaction With Servers}

Unlike Apache Thrift, Protocol Buffers do not include a concrete RPC
implementation.  However, serialized protocol buffers can trivially be
sent over TCP or integrated with a proprietary RPC system.  Combined
with an RPC system this means that one can interactively craft request
messages, send the serialized message to a remote server, read back a
response, and then parse the response protocol buffer interactively.

TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
Chu's Poly.

\section{Application: Protocol Buffers for Data Interchange in Web Services}

As the name implies, the primary application of protocol buffers is
data interchange in the context of inter-system communications. 
Network protocols such as HTTP describe procedures on client-server
communication, i.e. how to iniate requests, authenticate, send messages, 
etc. However, network 
protocols generally do not regulate \emph{content} of messages: they allow
transfer of any media type, such as web pages, files or video.
When designing systems where various components require exchange of specific data
structures, we need something on top of the protocol that prescribes 
how these structures are to be respresented in messages (buffers) on the
network. Protocol buffers solve exactly this problem by providing
a cross platform method for serializing arbitrary structures into well defined
messages, that can be exchanged using any protocol. The descriptors
(\texttt{.proto} files) are used to formally define the interface of a
remote API or network application. Libraries to parse and generate protobuf
messages are available for many programming languages, making it 
relatively straight forward to implement clients and servers.

@@

One example of a system that supports protocol buffers to interact
with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
computation and reproducible research based on R and Latex. It exposes a 
HTTP(S) API to access and manipulate R objects and allows for performing 
remote R function calls. Clients do not need to understand 
or generate any R code: HTTP requests are automatically mapped to 
function calls, and arguments/return values can be posted/retrieved
using several data interchange formats, such as protocol buffers.  
OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
from the \texttt{RProtoBuf} package to convert between R objects and protobuf
messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
earlier to parse and generate protobuf messages when interacting with OpenCPU.

@@

Because both HTTP and Protocol Buffers have libraries available for many 
languages, clients can be implemented in just a few lines of code. Below
example code for both R and Python that retrieve a dataset from R with 
OpenCPU using a protobuf message. In R, we use the HTTP client from 
the \texttt{httr} package \citep{httr}, and the protobuf
parser from the \texttt{RProtoBuf} package. In this illustrative example we
download a dataset which is part of the base R distribution, so we can actually
verify that the object was transferred without loss of information.

@@

# Check that no information was lost
identical(output, MASS::Animals)
@
This code suggests a method for exchanging objects between R servers, however this can 
also be done without protocol buffers. The main advantage of using an inter-operable format 
is that we can actually access R objects from within another
programming language. For example, in a very similar fasion we can retrieve the same
dataset in a Python client. To parse messages in Python, we first compile the 
\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:

@@

>>>>

\section{Application: Data Interchange in Web Services}

% TODO(jeroen): I think maybe some of this should go earlier in the
% paper, so this part can focus only on introducing the application,
% Can you integrate some of this text earlier, maybe into the the
% introduction?

As described earlier, the primary application of protocol buffers is
data interchange in the context of inter-system communications. 
Network protocols such as HTTP provide mechanisms for client-server
communication, i.e. how to initiate requests, authenticate, send messages, 
etc.  However, many network 
protocols generally do not regulate \emph{content} of messages: they allow
transfer of any media type, such as web pages, files or video.
When designing systems where various components require exchange of specific data
structures, we need something on top of the network protocol that prescribes 
how these structures are to be respresented in messages (buffers) on the
network. Protocol buffers solve exactly this problem by providing
a cross platform method for serializing arbitrary structures into well defined
messages, that can be exchanged using any protocol. The descriptors
(\texttt{.proto} files) are used to formally define the interface of a
remote API or network application. Libraries to parse and generate protobuf
messages are available for many programming languages, making it 
relatively straight forward to implement clients and servers.

@@

One example of a system that supports protocol buffers to interact
with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
computation and reproducible research based on R and \LaTeX. It exposes a 
HTTP(S) API to access and manipulate R objects and allows for performing 
remote R function calls. Clients do not need to understand 
or generate any R code: HTTP requests are automatically mapped to 
function calls, and arguments/return values can be posted/retrieved
using several data interchange formats, such as protocol buffers.  
OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
from the \texttt{RProtoBuf} package to convert between R objects and protobuf
messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
earlier to parse and generate protobuf messages when interacting with OpenCPU.

@@

Because both HTTP and Protocol Buffers have libraries available for many 
languages, clients can be implemented in just a few lines of code. Below
is example code for both R and Python that retrieves a dataset from R with 
OpenCPU using a protobuf message. In R, we use the HTTP client from 
the \texttt{httr} package \citep{httr}.
% superfluous?
%, and the protobuf parser from the \texttt{RProtoBuf} package.
In this example we
download a dataset which is part of the base R distribution, so we can
verify that the object was transferred without loss of information.

@@

# Check that no information was lost
identical(output, MASS::Animals)
@
This code suggests a method for exchanging objects between R servers, however this can 
also be done without protocol buffers. The main advantage of using an inter-operable format 
is that we can actually access R objects from within another
programming language. For example, in a very similar fashion we can retrieve the same
dataset in a Python client. To parse messages in Python, we first compile the 
\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:

@@

%\section{Application: Sending/receiving Interaction With Servers}
%
%Combined
%with an RPC system this means that one can interactively craft request
%messages, send the serialized message to a remote server, read back a
%response, and then parse the response protocol buffer interactively.

%TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
%Chu's Poly.
